---
title: "R Notebook"
output: html_notebook
---

###change

```{r}
library(leaflet)
library(readxl)
library(dplyr)
library(reticulate)
```


```{r}
maindata <- read_excel("maindata.xlsx")
tmp=maindata
```

Let look at how many unique values each column has

```{r}
sapply(maindata, function(x) length(unique(x)))
```

```{r}
maindata %>% group_by(cancellation_policy) %>% select(cancellation_policy)%>% unique()
```
```{r}
count(maindata, cancellation_policy) %>% mutate(relative_freq=(n/sum(n)))
```


```{r}
maindata %>% group_by(bed_type) %>% select(bed_type)%>% unique()
```

```{r}
library(dplyr)
maindata %>% select(property_type) %>% count(property_type) %>% mutate(relative_freq=(n/sum(n)))
```


```{r}
maindata %>% filter(city=="NYC") %>% select(bed_type) %>% count(bed_type) %>% mutate(relative_freq=(n/sum(n)))
```


```{r}
library(ggplot2)
maindata%>% ggplot( aes(x=cancellation_policy, y=log_price))+ geom_boxplot()+ ggtitle( "Boxplot for log_price vs cancellation policy")
```

```{r}
ggplot(maindata, aes(x=bed_type, y=log_price))+ geom_boxplot()+ ggtitle( "Boxplot for log_price vs bed type")
```

```{r}
ggplot(maindata, aes(x=cleaning_fee, y=log_price))+ geom_boxplot()+ ggtitle( "Boxplot for log_price vs cleaning fee")
```
Do some hypothesis testing.

```{r}
vcd::mosaic(~cleaning_fee+instant_bookable, data=maindata, shade=TRUE)
```


First, let us plot the different properties, using the latitude and longitude information given in our dataset. 
```{r}
m <- leaflet(tmp) %>%
addTiles() %>%
addProviderTiles("OpenStreetMap.BZH") %>%
addCircleMarkers(~long, ~lat, color = "red",
stroke=FALSE)
m
```
From the map above, we can see that the properties listed in our datset are from six different locations: Los Angeles, New York, DC, Boston, Chicago, San Fransico etc. 


## log_price for th different cities. 


### NYC 

```{r message=FALSE, warning=FALSE}

rows = (maindata$city== "NYC") 
tmp = maindata[rows, ]

mybins <- seq(2, 8, by=1.5)
mypalette <- colorBin( palette="YlOrBr",
domain=tmp$log_price,
na.color="transparent",
bins=mybins)

m <- leaflet(tmp) %>%
addTiles() %>%
addProviderTiles("OpenStreetMap.BZH") %>%
addCircleMarkers(~long, ~lat,radius = ~log_price,
fillColor = ~mypalette(log_price),
fillOpacity = 0.5,
color = "white",stroke=FALSE
)%>%
addLegend( pal=mypalette, values=~log_price, opacity=0.9,
title = "Log_price", position = "bottomright" )
m
```

Talk about which part of NYC has higher log_price and which part has lowest. 

### Boston.

```{r}

rows = (maindata$city== "Boston") 
tmp = maindata[rows, ]

mybins <- seq(2, 8, by=1.5)
mypalette <- colorBin( palette="YlOrBr",
domain=tmp$log_price,
na.color="transparent",
bins=mybins)

m <- leaflet(tmp) %>%
addTiles() %>%
addProviderTiles("OpenStreetMap.BZH") %>%
addCircleMarkers(~long, ~lat,radius = ~log_price,
fillColor = ~mypalette(log_price),
fillOpacity = 0.5,
color = "white",stroke=FALSE
)%>%
addLegend( pal=mypalette, values=~log_price, opacity=0.9,
title = "Log_price", position = "bottomright" )
m
```



### DC

```{r}

rows = (maindata$city== "DC") 
tmp = maindata[rows, ]

mybins <- seq(2, 8, by=1.5)
mypalette <- colorBin( palette="YlOrBr",
domain=tmp$log_price,
na.color="transparent",
bins=mybins)

m <- leaflet(tmp) %>%
addTiles() %>%
addProviderTiles("OpenStreetMap.BZH") %>%
addCircleMarkers(~long, ~lat,radius = ~log_price,
fillColor = ~mypalette(log_price),
fillOpacity = 0.5,
color = "white",stroke=FALSE
)%>%
addLegend( pal=mypalette, values=~log_price, opacity=0.9,
title = "Log_price", position = "bottomright" )
m
```


###  San Francisco. 

```{r}

rows = (maindata$city== "SF") 
tmp = maindata[rows, ]

mybins <- seq(2, 8, by=1.5)
mypalette <- colorBin( palette="YlOrBr",
domain=tmp$log_price,
na.color="transparent",
bins=mybins)

m <- leaflet(tmp) %>%
addTiles() %>%
addProviderTiles("OpenStreetMap.BZH") %>%
addCircleMarkers(~long, ~lat,radius = ~log_price,
fillColor = ~mypalette(log_price),
fillOpacity = 0.5,
color = "white",stroke=FALSE
)%>%
addLegend( pal=mypalette, values=~log_price, opacity=0.9,
title = "Log_price", position = "bottomright" )
m
```


### LA

```{r}

rows = (maindata$city== "LA") 
tmp = maindata[rows, ]

mybins <- seq(2, 8, by=1.5)
mypalette <- colorBin( palette="YlOrBr",
domain=tmp$log_price,
na.color="transparent",
bins=mybins)

m <- leaflet(tmp) %>%
addTiles() %>%
addProviderTiles("OpenStreetMap.BZH") %>%
addCircleMarkers(~long, ~lat,radius = ~log_price,
fillColor = ~mypalette(log_price),
fillOpacity = 0.5,
color = "white",stroke=FALSE
)%>%
addLegend( pal=mypalette, values=~log_price, opacity=0.9,
title = "Log_price", position = "bottomright" )
m
```


We can see that along the edge there are properties that are high priced. 


### Chicago

```{r}

rows = (maindata$city== "Chicago") 
tmp = maindata[rows, ]

mybins <- seq(2, 8, by=1.5)
mypalette <- colorBin( palette="YlOrBr",
domain=tmp$log_price,
na.color="transparent",
bins=mybins)

m <- leaflet(tmp) %>%
addTiles() %>%
addProviderTiles("OpenStreetMap.BZH") %>%
addCircleMarkers(~long, ~lat,radius = ~log_price,
fillColor = ~mypalette(log_price),
fillOpacity = 0.5,
color = "white",stroke=FALSE
)%>%
addLegend( pal=mypalette, values=~log_price, opacity=0.9,
title = "Log_price", position = "bottomright" )
m
```

## Types of property 

```{r}
rows = (maindata$city== "Chicago") 
tmp = maindata[rows, ]

mybins <- seq(2, 8, by=1.5)
mypalette <- colorBin( palette="YlOrBr",
domain=tmp$property_type,
na.color="transparent",
bins=mybins)

m <- leaflet(tmp) %>%
addTiles() %>%
addProviderTiles("OpenStreetMap.BZH") %>%
addCircleMarkers(~longitute, ~latitude,radius = ~log_price,
fillColor = ~mypalette(property_type),
fillOpacity = 0.5,
color = "white",stroke=FALSE
)%>%
addLegend( pal=mypalette, values=~property_type, opacity=0.9,
title = "Property type", position = "bottomright" )
m
max(maindata$log_price)
```

